%&../preamble

\def\npart {IA}
\def\nterm {Michaelmas}
\def\nyear {2021}
\def\nlecturer {R.\ Camina}
\def\ncourse {Groups}
\def\nauthor{}

\input{../preamble-dynamic.tex}

\includeonly{07-solids.tex}

\begin{document}
\maketitle

\section*{Syllabus}

{\small
  \noindent\textbf{Examples of groups}\\
  Axioms for groups. Examples from geometry: symmetry groups of regular polygons, cube, tetrahedron. Permutations on a set; the symmetric group. Subgroups and homomorphisms. Symmetry groups as subgroups of general permutation groups. The M\"obius group; cross-ratios, preservation of circles, the point at infinity. Conjugation. Fixed points of M\"obius maps and iteration.\hspace*{\fill} [4]

  \vspace{10pt}
  \noindent\textbf{Lagrange's theorem}\\
  Cosets. Lagrange's theorem. Groups of small order (up to order 8). Quaternions. Fermat-Euler theorem from the group-theoretic point of view.\hspace*{\fill} [5]

  \vspace{10pt}
  \noindent\textbf{Group actions}\\
  Group actions; orbits and stabilizers. Orbit-stabilizer theorem. Cayley's theorem (every group is isomorphic to a subgroup of a permutation group). Conjugacy classes. Cauchy's theorem.\hspace*{\fill} [4]

  \vspace{10pt}
  \noindent\textbf{Quotient groups}\\
  Normal subgroups, quotient groups and the isomorphism theorem.\hspace*{\fill} [4]

  \vspace{10pt}
  \noindent
  \textbf{Matrix groups}\\
  The general and special linear groups; relation with the M\"obius group. The orthogonal and special orthogonal groups. Proof (in $\mathbb{R}^3$) that every element of the orthogonal group is the product of reflections and every rotation in $\mathbb{R}^3$ has an axis. Basis change as an example of conjugation.\hspace*{\fill} [3]

  \vspace{10pt}
  \noindent\textbf{Permutations}\\
  Permutations, cycles and transpositions. The sign of a permutation. Conjugacy in $S_n$ and in $A_n$. Simple groups; simplicity of $A_5$.\hspace*{\fill} [4]}

\tableofcontents

\hypertarget{Introduction}{%
\section*{Introduction}\label{Introduction}}
\addcontentsline{toc}{section}{Introduction}

Dr Rachel Camina

Recommended books: Algebra \& Geometry, Alan Beardon.

Notation: \Lightning - contradiction

\newpage

\include{01.tex}
\include{02-dihedral.tex}
\include{03-cosets.tex}
\include{04-normal.tex}
\include{05-direct.tex}
\include{06-actions.tex}
\include{07-solids.tex}

\end{document}