\section{Transience and recurrence}
\subsection{Definitions}
\begin{definition}[Recurrent State]
	Let $X$ be a Markov chain, and let $i \in I$.
	A state $i$ is called \vocab{recurrent} if
	\begin{align*}
		\psub{i}{X_n = i \text{ for infinitely many } n} = 1
	\end{align*}
\end{definition}

\begin{definition}[Transient State]
	A state $i$ is called \vocab{transient} if
	\begin{align*}
		\psub{i}{X_n = i \text{ for infinitely many } n} = 0
	\end{align*}
\end{definition} 
We will prove that any $i$ is either recurrent or transient.

\subsection{Probability of visits}
\begin{definition}[Total number of visits]
	The number of visits to $i$ is defined by
	\begin{align*}
		V_i = \sum_{n=0}^\infty 1(X_n = i)\footnote{$1(X_0 = i) = 1$ so $V_i \geq 1$.}
	\end{align*}
\end{definition} 

\begin{definition}[Return Time]
	Let $T_i^{(0)} = 0$ and inductively define
	\begin{align*}
		T_i^{(r+1)} = \inf \qty{n \geq T_i^{(r)} + 1 \mid X_n = i}\footnote{First time we hit $i$ after $T_i^{(r)}$}
	\end{align*}
	We write $T_i^{(1)} = T_i$, called the first \vocab{return time} (or first passage time) to $i$.
	Further, let
	\begin{align*}
		f_i = \psub{i}{T_i^{(1)} < \infty}
	\end{align*}
\end{definition}

\begin{lemma}
	For all $r \in \mathbb N, i \in I$, $\psub{i}{V_i > r} = (f_i)^r$.
	So $V_i$ has a geometric distribution 
\end{lemma}
\begin{proof}
	For $r = 0$, this is trivially true.
	Now, suppose that the statement is true for $r$, and we will show that it is true for $r + 1$.
	\begin{align*}
		\psub{i}{V_i > r+1} & = \psub{i}{T_i^{(r+1)} < \infty}                                                      \\
		& = \psub{i}{T_i^{(r+1)} < \infty, T_i^{(r)} < \infty} \\
		& = \psub{i}{T_i^{(r+1)} < \infty \mid T_i^{(r)} < \infty} \psub{i}{T_i^{(r)} < \infty} \\
		& = \psub{i}{T_i^{(r+1)} < \infty \mid T_i^{(r)} < \infty} \psub{i}{V_i > r} \\
		& = \psub{i}{T_i^{(r+1)} < \infty \mid T_i^{(r)} < \infty} f_i^r 
		\intertext{The successive return times to $i$ are stopping times, so conditional on $T_i^{(k)} < \infty$ (and hence $T_i^{(k)} = i$) $\qty(X_{T_i^{(k)} + n})_{n \geq 0}$ is $\Markov{\delta_i, P}$, independent of $X_0, \dots, X_{T_i^{(k)}}$. So}
		% \intertext{By the strong Markov property applied to the stopping time $T_i^{(r)}$,}
		& = \psub{i}{T_i^{(1)} < \infty} f_i^r \\
		& = f_i f_i^r \\
		& = f_i^{r+1}
	\end{align*}
	
\end{proof}

\subsection{Duality of transience and recurrence}
\begin{theorem}
	Let $X$ be a Markov chain with transition matrix $P$, and let $i \in I$.
	Then, exactly one of the following is true.
	\begin{enumerate}
		\item If $f_i = 1$, then $i$ is recurrent, and
			\begin{align*}
				\sum_{n=0}^\infty p_{ii}(n) = \infty
			\end{align*}
		\item If $f_i < 1$, then $i$ is transient, and
			\begin{align*}
				\sum_{n=0}^\infty p_{ii}(n) < \infty
			\end{align*}
	\end{enumerate}
\end{theorem}

\begin{proof}
	\begin{align*}
		\esub{i}{V_i} & = \esub{i}{\sum_{n=0}^\infty 1(X_n = i)} \\
		& = \sum_{n=0}^\infty \esub{i}{1(X_n = i)} \\
		& = \sum_{n=0}^\infty \psub{i}{X_n = i}    \\
		& = \sum_{n=0}^\infty p_{ii}(n)
	\end{align*}
	First, suppose $f_i = 1$.
	Then, for all $r$, $\psub{i}{V_i > r} = 1$, so $\psub{i}{V_i = \infty} = 1$.
	Hence, $i$ is recurrent.
	Further, $\esub{i}{V_i} = \infty$ so $\sum_{n=0}^\infty p_{ii}(n) = \infty$.

	Now, if $f_i < 1$, by the previous lemma we see that $\esub{i}{V_i} = \frac{1}{1-f_i} < \infty$ hence $\psub{i}{V_i < \infty} = 1$.
	Thus, $i$ is transient.
	Further, $\esub{i}{V_i} < \infty$ so $\sum_{n=0}^\infty p_{ii}(n) < \infty$.
\end{proof}
\begin{theorem}
	Let $x, y$ be states that communicate.
	Then, either $x$ and $y$ are both recurrent, or they are both transient.
\end{theorem}
\begin{proof}
	Suppose $x$ is recurrent.
	Then, since $x$ and $y$ communicate, $\exists m, \ell \in \mathbb N$ such that
	\begin{align*}
		p_{xy}(m) > 0;\quad p_{yx}(\ell) > 0
	\end{align*}
	Note, $\sum_n p_{xx}(n) = \infty$.
	Then,
	\begin{align*}
		\sum_n p_{yy}(n) &\geq \sum_n p_{yy}(n+m+\ell) \geq \sum_n p_{yx}(\ell) p_{xx}(n) p_{xy}(m) \\
		&\geq p_{yx}(\ell) p_{xy}(m) \sum_n p_{xx}(n) = \infty
	\end{align*}
\end{proof}
\begin{corollary}
	Either all states in a communicating class are recurrent or they are all transient.
\end{corollary}

\subsection{Recurrent communicating classes}
\begin{theorem}
	Any recurrent communicating class is closed.
\end{theorem}
\begin{proof}
	Suppose a communicating class $C$ is not closed.
	Then there exists $x \in C$ and $y \not\in C$ such that $x \to y$.
	Let $m$ be such that $p_{xy}(m) > 0$.
	If, starting from $x$, we hit $y$ which is outside the communicating class, then we can never return to the communicating class (including $x$) again.
	In particular,
	\begin{align*}
		\psub{x}{V_x < \infty} \geq \psub{x}{X_m = y} = p_{xy}(m) > 0
	\end{align*}
	Hence $x$ is not recurrent, which is a contradiction.
\end{proof}
\begin{theorem}
	Any finite closed communicating class is recurrent.
\end{theorem}
\begin{proof}
	Let $C$ be a finite closed communicating class.
	Let $x \in C$.
	Then, by the pigeonhole principle, there must exist $y \in C$ such that
	\begin{align*}
		\psubx{X_n = y \text{ for infinitely many } n} > 0
	\end{align*}
	Since $x$ and $y$ communicate, there exists $m \in \mathbb N$ such that $p_{yx}(m) > 0$.
	Now,
	\begin{align*}
		\psub{y}{X_m = y \text{ for infinitely many } n} & \geq \psubx{X_m = x, X_n = y \text{ for infinitely many } n \geq m}                   \\
		                                                 & =\psubx{X_n = y \text{ for infinitely many } n \geq m \mid X_m = x} \psub{y}{X_m = x} \\
		                                                 & =\psubx{X_n = y \text{ for infinitely many } n} \psub{y}{X_m = x} > 0
	\end{align*}
	Thus $y$ is recurrent.
	Since recurrence is a class property, $C$ is recurrent.
\end{proof}
\begin{theorem}
	Let $P$ be irreducible and recurrent.
	Then, for all $x, y$,
	\begin{align*}
		\psub{x}{T_y < \infty} = 1
	\end{align*}
\end{theorem}
\begin{proof}
	Since $y$ is recurrent,
	\begin{align*}
		1 = \psub{y}{X_n = y \text{ for infinitely many } n}
	\end{align*}
	Let $m$ such that $p_{yx}(m) > 0$.
	Now,
	\begin{align*}
		1 & = \psub{y}{X_n = y \text{ infinitely often}}                                                       \\
		  & = \sum_z \psub{y}{X_m = z, X_n = y \text{ for infinitely many } n \geq m}                          \\
		  & = \sum_z \psub{y}{X_n = y \text{ for infinitely many } n \geq m \mid X_m = z} \psub{y}{X_m = z}    \\
		  & = \sum_z \psub{z}{X_n = y \text{ for infinitely many } n} \psub{y}{X_m = z}                        \\
		\intertext{By the strong Markov property,}
		  & = \sum_z \psub{z}{T_y < \infty} \psub{y}{X_n = y \text{ for infinitely many } n} \psub{y}{X_m = z} \\
		\intertext{Since $y$ is recurrent,}
		  & = \sum_z \psub{z}{T_y < \infty} \psub{y}{X_m = z}                                                  \\
		  & = \sum_z \psub{z}{T_y < \infty} p_{yz}(m)
	\end{align*}
	Since $p_{yz}(m) > 0$ and $\sum_z p_{yz}(m) = 1$, $\psub{x}{T_y < \infty} = 1$.
\end{proof}