\section{$p$-groups}

\subsection{$p$-groups}
\begin{definition}[$p$-group]
	Let $p$ be a prime.
	A finite group $G$ is a \vocab{$p$-group} if $\abs{G} = p^n$ for $n \geq 1$.
\end{definition}

\begin{theorem}
	If $G$ is a $p$-group, the centre $Z(G)$ is non-trivial.
\end{theorem}

\begin{proof}
	For $g \in G$, due to the orbit-stabiliser theorem, $\abs{\ccl(g)} \abs{C(g)} = p^n$.
	In particular, $\abs{\ccl(g)}$ divides $p^n$, and they partition $G$.
	Since $G$ is a disjoint union of conjugacy classes, modulo $p$ we have
	\begin{align*}
		\abs{G} \equiv \text{number of conjugacy classes of size } 1 \equiv 0 \implies \abs{Z(G)} \equiv 0
	\end{align*}
	Hence $Z(G)$ has order zero modulo $p$ so it cannot be trivial.
	We can check this by noting that $g \in Z(G) \iff x\inv g x = g$ for all $x$, which is true if and only if $\ccl_G(g) = \qty{g}$.
\end{proof}

\begin{corollary}
	The only simple $p$-groups are the cyclic groups of order $p$.
\end{corollary}

\begin{proof}
	Let $G$ be a simple $p$-group.
	Since $Z(G)$ is a normal subgroup of $G$, we have $Z(G) = 1$ or $Z(G) = G$.
	But $Z(G)$ may not be trivial, so $Z(G) = G$.
	This implies $G$ is abelian.
	The only abelian simple groups are cyclic of prime order by \cref{lem:1.3}, hence $G \cong C_p$.
\end{proof}

\begin{corollary}
	Let $G$ be a $p$-group of order $p^n$.
	Then $G$ has a subgroup of order $p^r$ for all $r \in \qty{0, \dots, n}$.
\end{corollary}

\begin{proof}
	Recall from \cref{lem:1.4} that any group $G$ has a composition series $1 = G_1 \triangleleft \dots \triangleleft G_N = G$ where each quotient $\faktor{G_{i+1}}{G_i}$ is simple.

	Since $G$ is a $p$-group, $\faktor{G_{i+1}}{G_i}$ is also a $p$-group.
	Each successive quotient is an order $p$ group by the previous corollary, so we have a composition series of nested subgroups of order $p^r$ for all $r \in \qty{0, \dots, n}$.
\end{proof}

\begin{lemma}
	Let $G$ be a group.
	If $\faktor{G}{Z(G)}$ is cyclic, then $G$ is abelian.
	This then implies that $Z(G) = G$, so in particular $\faktor{G}{Z(G)} = 1$.
\end{lemma}

\begin{proof}
	Let $g Z(G)$ be a generator for $\faktor{G}{Z(G)}$.
	Then, each coset of $Z(G)$ in $G$ is of the form $g^r Z(G)$ for some $r \in \mathbb Z$.
	Thus, $G = \qty{g^r z \colon r \in \mathbb Z, z \in Z(G)}$.
	Now, we multiply two elements of this group and find
	\begin{align*}
		g^{r_1} z_1 g^{r_2} z_2 = g^{r_1 + r_2} z_1 z_2 = g^{r_1 + r_2} z_2 z_1 = z_2 z_1 g^{r_1 + r_2} = g^{r_2} z_2 g^{r_1} z_1
	\end{align*}
	So any two elements in $G$ commute.
\end{proof}

\begin{corollary}
	Any group of order $p^2$ is abelian.
\end{corollary}

\begin{proof}
	Let $G$ be a group of order $p^2$.
	Then $\abs{Z(G)} \in \qty{1, p, p^2}$.
	The centre cannot be trivial as proven above, since $G$ is a $p$-group.
	If $\abs{Z(G)} = p$, we have that $\faktor{G}{Z(G)}$ is cyclic as it has order $p$.
	Applying the previous lemma, $G$ is abelian.
	However, this is a contradiction since the centre of an abelian group is the group itself.
	If $\abs{Z(G)} = p^2$ then $Z(G) = G$ and then $G$ is clearly abelian.
\end{proof}

\subsection{Sylow theorems}
\begin{theorem}[Sylow Theorems]
	Let $G$ be a finite group of order $p^a m$ where $p$ is a prime and $p$ does not divide $m$.
	Then:
	\begin{enumerate}
		\item The set $\mathrm{Syl}_p(G) = \qty{P \leq G \colon \abs{P} = p^a}$ of Sylow $p$-subgroups is non-empty.
		\item All Sylow $p$-subgroups are conjugate.
		\item The amount of Sylow $p$-subgroups $n_p = \abs{\mathrm{Syl}_p(G)}$ satisfies
		      \begin{align*}
			      n_p \equiv 1 \mod p;\quad n_p \mid \abs{G} \implies n_p \mid m
		      \end{align*}
	\end{enumerate}
\end{theorem}

\begin{proof}
	\begin{enumerate}
		\item Let $\Omega$ be the set of all \underline{subsets} of $G$ of order $p^a$.
		      We can directly find
		      \begin{align*}
			      \abs{\Omega} = \binom{p^a m}{p^a} = \frac{p^a m}{p^a} \cdot \frac{p^a m - 1}{p^a - 1} \cdots \frac{p^a m - p^a + 1}{1}
		      \end{align*}
		      Note that for $0 \leq k < p^a$, the numbers $p^a m - k$ and $p^a - k$ are divisible by the same power of $p$.
		      In particular, $\abs{\Omega}$ is coprime to $p$.

		      Let $G$ act on $\Omega$ by left-multiplication, so $g \ast X = \qty{gx \colon x \in X}$.
		      For any $X \in \Omega$, the orbit-stabiliser theorem can be applied to show that
		      \begin{align*}
			      \abs{G_X} \abs{\mathrm{orb}_G(X)} = \abs{G} = p^a m
		      \end{align*}
		      Since $|\Omega|$ is coprime to $p$, there must exist an orbit with size coprime to $p$, since orbits partition $\Omega$.
		      For such an $X$, $p^a \mid \abs{G_X}$.

		      Conversely, note that if $g \in G$ and $x \in X$, then $g \in (gx\inv) \ast X$.
		      Hence, we can consider
		      \begin{align*}
			      G = \bigcup_{g \in G} g \ast X = \bigcup_{Y \in \mathrm{orb}_G(X)} Y
		      \end{align*}
		      Thus $\abs{G} \leq \abs{\mathrm{orb}_G(X)} \cdot \abs{X}$, giving $\abs{G_X} = \frac{\abs{G}}{\abs{\mathrm{orb}_G(X)}} \leq \abs{X} = p^a$.

		      As $p^a \mid \abs{G_X}$ we must have $\abs{G_X} = p^a$.
		      In other words, the stabiliser $G_X$ is a Sylow $p$-subgroup of $G$.
		\item We will prove a stronger result for this part of the proof.
		      \begin{lemma} \label{lem:4.7}
				If $P$ is a Sylow $p$-subgroup and $Q \leq G$ is a $p$-subgroup, then $Q \leq g P g\inv$ for some $g \in G$.
			  \end{lemma} 

		      Indeed, let $Q$ act on the set of left cosets of $P$ in $G$ by left multiplication.
		      By the orbit-stabiliser theorem, each orbit has size which divides $\abs{Q} = p^k$ for some $k$.
		      Hence each orbit has size $p^r$ for some $r$.

		      Since $\faktor{G}{P}$ has size $m$, which is coprime to $p$, there must exist an orbit of size 1\footnote{Sum of the orbit sizes is $m$, $m$ coprime to $p$.}.
		      Therefore there exists $g \in G$ such that $q \ast gP = gP$ for all $q \in Q$.
		      Equivalently, $g\inv q g \in P$ for all $q \in Q$.
		      This implies that $Q \leq gPg\inv$ as required.
		      This then weakens to the second part of the Sylow theorems.
		\item Let $G$ act on $\mathrm{Syl}_p(G)$ by conjugation.
		      Part (ii) of the Sylow theorems implies that this action is transitive.
		      By the orbit-stabiliser theorem, $n_p = \abs{\mathrm{Syl}_p(G)} \mid \abs{G}$.

		      Let $P \in \mathrm{Syl}_p(G)$.
		      Then let $P$ act on $\mathrm{Syl}_p(G)$ by conjugation.
		      Since $P$ is a Sylow $p$-subgroup, the orbits of this action have size dividing $\abs{P} = p^a$, so the size is some power of $p$. \\
		      To show $n_p \equiv 1 \mod p$, it suffices to show that $\qty{P}$ is the unique orbit of size 1, as the orbits of other sizes are multiples of $p$. \\
		      Suppose $\qty{Q}$ is another orbit of size 1, so $Q$ is a Sylow $p$-subgroup which is preserved under conjugation by $P$.
		      Thus $P$ normalises $Q$, so $P \leq N_G(Q)$.
		      Notice that $P$ and $Q$ are both Sylow $p$-subgroups of $N_G(Q)$.
		      By (ii), $P$ and $Q$ are conjugate inside $N_G(Q)$.
		      Hence $P = Q$ since $Q \trianglelefteq N_G(Q)$.
		      Thus, $\abs{P}$ is the unique orbit of size 1, so $n_p \equiv 1 \mod p$ as required.
	\end{enumerate}
\end{proof}

\begin{corollary}
	If $n_p = 1$, then there is only one Sylow $p$-subgroup, and it is normal.
\end{corollary}

\begin{proof}
	Let $g \in G$ and $P \in \mathrm{Syl}_p(G)$.
	Then $g P g\inv$ is a Sylow $p$-subgroup, hence $g P g\inv = P$.
	$P$ is normal in $G$.
\end{proof}

\begin{remark}
	When $G$ acts on $\mathrm{Syl}_p(G)$ by conjugation, the orbit is $\mathrm{Syl}_p(G)$ and the stabiliser is the normaliser.  
\end{remark}

\begin{example}
	Let $G$ be a group with $\abs{G} = 1000 = 2^3 \cdot 5^3$.
	Here, $n_5 \equiv 1 \mod 5$, and $n_5 \mid 8$, hence $n_5 = 1$.
	Thus the unique Sylow 5-subgroup is normal.
	Hence no group of order 1000 is simple.
\end{example}

\begin{example}
	Let $G$ be a group with $\abs{G} = 132 = 2^2 \cdot 3 \cdot 11$.
	$n_{11}$ satisfies $n_{11} \equiv 1 \mod 11$ and $n_{11} \mid 12$, thus $n_{11} \in \qty{1, 12}$.

	Suppose $G$ is simple.

	Then $n_{11} = 12$\footnote{If $n_{11} = 1$ then we have a normal subgroup by the previous corollary.}.
	The amount of Sylow 3-subgroups satisfies $n_3 \equiv 1 \mod 3$ and $n_3 \mid 44$ so $n_3 \in \qty{1, 4, 22}$.
	Since $G$ is simple, $n_3 \in \qty{4, 22}$.

	Suppose $n_3 = 4$.
	Then $G$ acts on $\mathrm{Syl}_3(G)$ by conjugation, and this generates a group homomorphism $\varphi \colon G \to S_4$.
	But the kernel of this homomorphism is a normal subgroup of $G$, so $\ker \varphi$ is trivial or $G$ itself as $G$ simple.
	If $\ker \varphi = G$, then $\Im \varphi$ is trivial, contradicting Sylow's second theorem.
	If $\ker \varphi = 1$, then $\Im \varphi$ has order $132 > |S_4|$ \Lightning.

	Thus $n_3 = 22$ and recall $n_{11} = 12$.
	This means that $G$ has $22 \cdot (3-1) = 44$ elements of order 3\footnote{Each group in $\mathrm{Syl}_3(G)$ intersect trivially, as if they didn't any non trivial element in the intersection would generate both groups as they're all $C_3$.}, and further $G$ has $12 \cdot (11 - 1) = 120$ elements of order 11.
	However, the sum of these two totals is more than the total of 132 elements, so this is a contradiction.
	Hence $G$ is not simple.
\end{example}
