%&../preamble

\def\npart {II}
\def\nterm {Michaelmas}
\def\nyear {2023}
\def\nlecturer {Dr Sarkar}
\def\ncourse {Probability and Measure}

\input{../preamble-dynamic.tex}

% \includeonly{07_ergodic_theory.tex}

\setcounter{section}{-1}

\begin{document}
    \maketitle
    \tableofcontents

    \section{Holes in classical theory}

    Analysis

    \begin{enumerate}
        \item What is the ``volume'' of a subset of $\mathbb{R}^d$.
        \item Integration (Riemann Integration has holes)
        \begin{itemize}
            \item $\{f_n\}$ a sequence of continuous functions on $[0, 1]$ s.t.
            \begin{itemize}
                \item $0 \leq f_n(x) \leq 1 \; \forall \; x \in [0, 1]$.
                \item $f_n(x)$ is monotonically decreasing on $n \to \infty$, i.e. $f_n(x) \geq f_{n+1}(x) \; \forall \; x$/
            \end{itemize}
            So, $\lim_{n \to \infty} f_n(x)$ exists. But $f$ is not Riemann integrable . We want a theory of integration s.t. $f$ is integrable and $\lim_{n \to \infty} \int_{0}^{1} f_n(x) \dd{x} = \int_{0}^{1} f(x) \dd{x}$.
        \end{itemize}
        \item $L^1 = ()$
        If $f \in L^1$ is $f$ Riemann integrable? Will have to change the definition of integral. $L^2$ a hilbert space
    \end{enumerate}

    Probability

    \begin{enumerate}
        \item Discrete probability has its limitations,
        \begin{itemize}
            \item Toss a unbiased coin 5 times. What is the probability if getting 3 heads?
            \item Take an infinite sequence of coin tosses ($E = \{0, 1\}^\mathbb{N}$ which is uncountable) and an event A that depends on that infinite sequence. How do you define $\mathbb{P}(A)$?
            E.g. $X_i \sim \operatorname{Ber}\left( \frac{1}{2} \right)$ and $A = \frac{\sum_{i=1}^{n} X_i}{n}$, the average number of heads.
            By strong law of large numbers $\mathbb{P}\left( \frac{\sum_{i=1}^{n} X_i}{n} \to \frac{1}{2} \right) = 1$.
            \item How to draw a point uniformly at random from $[0, 1]$? $U \sim U[0, 1]$.
            Probability needs axioms to be made rigorous.
        \end{itemize}
        \item Define Expectation for a r.v.. Also would want the following if $0 \leq X_n \leq 1$ and $X_n \downarrow X$ then $\mathbb{E} X_n \to \mathbb{E} X$.
    \end{enumerate}

    \include{01_measures.tex}
    \include{02_measurable_functions.tex}
    \include{03_integration.tex}
    \include{04_product_measures.tex}
    \include{05_function_spaces_and_norms.tex}
    \include{06_fourier_analysis.tex}
    \include{07_ergodic_theory.tex}
\end{document}